---
title: "General Social Survey"
author: "Monique Sisneros"
date: "April 27,2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

mydata5 <- filter(mydata4, CAPPUN=="FAVOR" | CAPPUN=="OPPOSE")

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. First explain what you make out of comparing how people voted in 2012 with 2016.


How are the "voted in 2016" and "voted in 2012 related?
$H_0:$ voting in 2016 and 2012 are unrelated
The alternative hypothesis: is voting in 2016 directly related to the way people voted in 2012.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->


The methods used were categorical variables and a chi- squared test to test the hypothesis that voting in 2016 is / is not independent of voting in 2012.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->


### Descriptive Results
We use a barchart to show the voting trends between 2012 and 2016.



#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent") 

```
#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Numerical summary is a cross tabulation:
```{r}
xtabs(~VOTE16+VOTE12, data=mydata4)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```
# Discussion and Conclusion

The hypothesis was that voting in 2012 is 2016 related . We used a chi square test of indepenednce the p vaule is equal to 0. So therefore the null hypothesis is related.
We reject the null hypothesis because the p vaule is less then .05. We have evidence that voting in 2012 is related to voting in 2016 and if you did not vote in 2012 you likely did not vote iin 2016.

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

How are the "Big Bang Theory" and "belief in the after life" related?
$H_0:$ If people believe in the big bang theory also beleieve in the after life.
Alternative hypothesis: There is no relation between people who believe in the big bang theory and people who believe in an after life.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!--> 
We used a graphical chart,bar chart,chi chart,and a summary x tab.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata3, type = "percent", xlab='BELIEVE IN POSTLIFE', main='BELIEVE IN BIGBANG')
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Numerical summary is a cross tabulation:

```{r}
xtabs(~BIGBANG+POSTLIFE, data=mydata3)
```





### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata3)
BANGtab=xtabs(~BIGBANG+POSTLIFE, data=mydata3)
fisher.test(BANGtab)
```
The two concept we are looking at is if the people believe in the bigbang theory and beleieve in the post life. The null hypothesis has no relations between each other. So there is not relation between the two.

# Discussion and Conclusion
The hypothesis was that the belief in the big bang theory is related to the belief in the after life. We used a chi square test of independence. because the p vaule is less then .05. Meaning the null hypothesis is related. We rejected the null since we have evidence that the belief in the postlife is related to the belief in the big bang theory and if you do not believe in the after life you likely did not believe in the big bang theory. 
# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This time compare the opinion on death penalty (CAPPUN) with opinion on courts (COURTS).
How are the "CAPPUN" and "COURTS" related?
$H_0:$ If people believe in the big bang theory also beleieve in the after life.
Alternative hypothesis: There is no relation between people who believe death penalty  and people who believe in the courts.
## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We used a graphical chart,bar chart,chi chart,and a summary x tab.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+COURTS, data=mydata5, type = "percent", xlab='View on CAPPUN', main='View on COURTS')
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Numerical summary is a cross tabulation:
```{r}
xtabs(~CAPPUN+COURTS, data=mydata5)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~CAPPUN+COURTS, data=mydata5)
CAPPUNtab=xtabs(~CAPPUN+COURTS, data=mydata5)
fisher.test(CAPPUNtab)
```


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The hypothesis was that do people believe in the death penalty is related to the belief in the court system. We used a chi square test of independence. because the p vaule is less then .05. Meaning the null hypothesis is related. We rejected the null since we have evidence that the belief in the death penalty is related to the belief in the court system and if you do not believe in the death penalty it is likely you did not believe in the court system.

